LUCID: Learning-Enabled Uncertainty-Aware Certification of Stochastic Dynamical Systems

### 2025-07-30 docs/paper/main.tex
feat: introduce barrier 2

\todo{Platform, configuration, seed for reproducibility}

@@

@@

The \barr experiment is inspired by the one from \cite{abate2021fossil}, which we extend by adding stochastic noise $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$ to arrive at the two-dimensional nonlinear stochastic dynamics
\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{2, t} \\
        \frac{1}{3} {x}^3_{1, t} - {x}_{1,t} - {x}_{2,t}
    \end{bmatrix} + w_t.
\end{equation*}
Given
\begin{align*}
    &\X = [ -3, 2.5 ] \times [ -2 , 1 ] \\
    &\X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ] \\
    &\qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1] \\
    &\X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=10$ time steps.
\EC{we can remove all the explicit set intevals}
This example is particularly challenging due to the closeness between the initial set and the unsafe set.
Hence, we need a rather high number of frequencies to achieve the required expressivity in the barrier.
The results are shown in Figure~\ref{fig:CSBarr3}.
The kernel \hp have been selected through offline tuning to be $\sigma_f=X$ and $\sigma_l=[X, X]$.
Using a spectral basis of $M=X^3-1$ frequencies and a lattice density of $\hat{N} = X^2$, 
we were able to generate a \gls{cbc} with $\eta \approx X.X$, $\gamma \approx X.X$, and $c \approx X.X$ in $X$ minutes and $X$ seconds, achieving a safety probability of $P_{safe}(\S^\pi) \approx XX.X\%$.
% Knowing the system dynamics, we were also able to formally validate the barrier.

\begin{figure}[ht]
    \resizebox{\columnwidth}{!}{\import{figures}{barrier3.pgf}}%
    \caption{\gls{cbc} synthesis for the \barr benchmark.
        The blue regions represent the initial set $\X_0$, while the red regions represent the unsafe set $\X_U$.
        The pink surface indicates the value of $\gamma$.
    }
    \label{fig:CSBarr3}
\end{figure}

>>>>

@@

\subsection{Barrier 2}

The \barrII experiment is inspired by the one from \cite{abate2021fossil}, which we extend by adding stochastic noise $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$ to arrive at the two-dimensional nonlinear stochastic dynamics

\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{2, t} - 1 + e^{-x_{1, t}} \\
        -\sin^2(x_{1, t})
    \end{bmatrix} + w_t.
\end{equation*}
Given
\begin{align*}
    &\X = [ -2, 2 ] \times [ -2 , 2 ] \\
    &\X_0 = \{ [x_1, x_2] : (x_1 + 0.5)^2 + (x_2 - 0.5)^2 \leq 0.4 \} \\
    &\X_u = \{ [x_1, x_2] : (x_1 - 0.7)^2 + (x_2 + 0.7)^2 \leq 0.3 \}
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
\EC{we can remove all the explicit set intevals}
The results are shown in Figure~\ref{fig:CSBarr2}.
The kernel \hp have been selected through offline tuning to be $\sigma_f=X$ and $\sigma_l=[X, X]$.
Using a spectral basis of $M=X^3-1$ frequencies and a lattice density of $\hat{N} = X^2$, 
we were able to generate a \gls{cbc} with $\eta \approx X.X$, $\gamma \approx X.X$, and $c \approx X.X$ in $X$ minutes and $X$ seconds, achieving a safety probability of $P_{safe}(\S^\pi) \approx XX.X\%$.
% Knowing the system dynamics, we were also able to formally validate the barrier.

\begin{figure}[ht]
    \resizebox{\columnwidth}{!}{\import{figures}{barrier3.pgf}}%
    \caption{\gls{cbc} synthesis for the \barrIII benchmark.
        The blue regions represent the initial set $\X_0$, while the red regions represent the unsafe set $\X_U$.
        The pink surface indicates the value of $\gamma$.
    }
    \label{fig:CSBarr2}
\end{figure}

@@

The \barrIII experiment is inspired by the one from \cite{abate2021fossil}, which we extend by adding stochastic noise $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$ to arrive at the two-dimensional nonlinear stochastic dynamics
\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{2, t} \\
        \frac{1}{3} {x}^3_{1, t} - {x}_{1,t} - {x}_{2,t}
    \end{bmatrix} + w_t.
\end{equation*}
Given
\begin{align*}
    &\X = [ -3, 2.5 ] \times [ -2 , 1 ] \\
    &\X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ] \\
    &\qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1] \\
    &\X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=10$ time steps.
\EC{we can remove all the explicit set intevals}
This example is particularly challenging due to the closeness between the initial set and the unsafe set.
Hence, we need a rather high number of frequencies to achieve the required expressivity in the barrier.
The results are shown in Figure~\ref{fig:CSBarr3}.
The kernel \hp have been selected through offline tuning to be $\sigma_f=X$ and $\sigma_l=[X, X]$.
Using a spectral basis of $M=X^3-1$ frequencies and a lattice density of $\hat{N} = X^2$, 
we were able to generate a \gls{cbc} with $\eta \approx X.X$, $\gamma \approx X.X$, and $c \approx X.X$ in $X$ minutes and $X$ seconds, achieving a safety probability of $P_{safe}(\S^\pi) \approx XX.X\%$.
% Knowing the system dynamics, we were also able to formally validate the barrier.

\begin{figure}[ht]
    \resizebox{\columnwidth}{!}{\import{figures}{barrier3.pgf}}%
    \caption{\gls{cbc} synthesis for the \barrIII benchmark.
        The blue regions represent the initial set $\X_0$, while the red regions represent the unsafe set $\X_U$.
        The pink surface indicates the value of $\gamma$.
    }
    \label{fig:CSBarr3}
\end{figure}
